Problem of Step 4 (debias) of fGFPCA on NHANES data

### 2024-02-28 Docs/NhanesApplExample.Rmd
more explorations on the reversed eigenvalue problem

```{r}
# debias model
gen_df <- gen_df %>% 
  left_join(df_phi, by = "sind")
gen_df$id <- as.factor(gen_df$id)
debias_glmm2 <- bam(Y ~ s(sind, bs="cr")+
                   s(id, by=phi1, bs="re")+
                   s(id, by=phi2, bs="re")+
                   s(id, by=phi3, bs="re")+
                   s(id, by=phi4, bs="re"),
                 family = binomial,
                 data=gen_df,
                 method = "fREML",
                 discrete = TRUE)
```

@@

```{r, eval=FALSE}
new_df <- gen_df %>% filter(id == 1)
est_phi <- predict(debias_glmm2, newdata = new_df, type = "terms")[, 2:5] %>% 
  data.frame()
est_phi %>%
  mutate(sind=1:J) %>%
  pivot_longer(1:4) %>%
  ggplot()+
  geom_line(aes(x=sind, y= value))+
  facet_wrap(~name)
```

```{r eval=FALSE}
coef1 <- coef(debias_glmm2)[c("s(id):phi1.1", "s(id):phi2.1", "s(id):phi3.1", "s(id):phi4.1")]
coef1 <- matrix(coef1, nrow = J, ncol = 4, byrow = T)
(as.matrix(df_phi[, 2:5]) * coef1) %>% data.frame() %>%
  mutate(sind=1:J) %>%
  pivot_longer(1:4) %>%
  ggplot()+
  geom_line(aes(x=sind, y= value))+
  facet_wrap(~name)

@@

>>>>

```{r}
# debias model
gen_df <- gen_df %>% 
  left_join(df_phi, by = "sind")
gen_df$id <- as.factor(gen_df$id)
debias_glmm2 <- bam(Y ~ s(sind, bs="cr")+
                   s(id, by=phi1, bs="re")+
                   s(id, by=phi2, bs="re")+
                   s(id, by=phi3, bs="re")+
                   s(id, by=phi4, bs="re"),
                 family = binomial,
                 data=gen_df,
                 method = "fREML",
                 discrete = TRUE,
                 use.chol = TRUE)
```

@@

@@

# Other random explorations

- what if I fit the bam model in a subset of data (e.g. bin midpoints) ?

```{r}
try_glmm <- bam(Y ~ s(sind, bs="cr")+
                   s(id, by=phi1, bs="re")+
                   s(id, by=phi2, bs="re")+
                   s(id, by=phi3, bs="re")+
                   s(id, by=phi4, bs="re"),
                 family = binomial,
                 data=gen_df %>% filter(sind==bin),
                 method = "fREML",
                 discrete = TRUE,
                 use.chol = TRUE)

1/try_glmm$sp[2:5]

- What about sub-intervals of time (e.g. every three hours? )

```{r}
evals_by_hour <- gen_df %>% 
  mutate(hour = cut(sind, breaks = seq(0, J, by = 180), labels = 1:8)) %>%
  group_by(hour) %>%
  # filter(hour %in% 1:3) %>%
  group_map(~{
    try_glmm <- bam(Y ~ s(sind, bs="cr")+
                   s(id, by=phi1, bs="re")+
                   s(id, by=phi2, bs="re")+
                   s(id, by=phi3, bs="re")+
                   s(id, by=phi4, bs="re"),
                 family = binomial,
                 data=.x,
                 method = "fREML",
                 discrete = TRUE)
    1/try_glmm$sp[2:5]
  })

bind_rows(evals_by_hour) %>%
  mutate(hour = seq(3, 24, by = 3)) %>% 
  pivot_longer(1:4) %>% 
  ggplot(aes(x=hour, y=value, group=name, col=name))+
  geom_point()+
  labs(x="Hour", y= "Evals")
```
